---
title: "Figure 3"
output: html_document
date: "2026-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Figure 3A - Linear regression

##1. diet study
###1.1 Preparation
```{r}
vars_to_normalize <- c("Fecal_Mass", "Microbiota_Excretion", "Microbial_Load", "Transit_Time")

QMPs_ileum_diet_fold <- diet_data_ileum %>%
  group_by(project) %>%
  mutate(across(
    all_of(vars_to_normalize),
    ~ .x / mean(.x[diet == "control"], na.rm = TRUE)
  )) %>%
  ungroup()

QMPs_colon_diet_fold <- diet_data_colon %>%
  group_by(project) %>%
  mutate(across(
    all_of(vars_to_normalize),
    ~ .x / mean(.x[diet == "control"], na.rm = TRUE)
  )) %>%
  ungroup()
```

###1.2 Linear Regression
```{r}

variables <- c("Fecal_Mass", "Microbiota_Excretion", "Microbial_Load", "Transit_Time")

effect_results_diet <- list()
emmeans_results_diet <- list()

for (var in variables) {
  lm_data <- QMPs_ileum_diet_fold %>%
    mutate(
      logY = log10(.data[[var]]),
      diet = factor(diet)
    )
  fit <- lm(logY ~ diet, data = lm_data)
  
  res <- tidy(fit) %>%
    filter(term != "(Intercept)") %>%
    mutate(variable = var)
  effect_results_diet[[var]] <- res
  
  emm <- emmeans(fit, ~ diet)
  
  pairs_res <- pairs(emm) %>% as.data.frame() %>% mutate(variable = var)
  emmeans_results_diet[[var]] <- pairs_res
}

effect_results_diet <- bind_rows(effect_results_diet)
emmeans_results_diet <- bind_rows(emmeans_results_diet)

effect_results_diet <- effect_results_diet %>%
  group_by(variable) %>%
  mutate(FDR = p.adjust(p.value, method = "fdr")) %>%
  ungroup()

```

### 3.1 Plot
```{r}
plot_df_diet <- effect_results_diet %>%
  mutate(
    lower.CL = estimate - 1.96 * std.error,
    upper.CL = estimate + 1.96 * std.error
  )

diet_labels <- c(
  "dietfasting"    = "Fasting",
  "diethigh_fat"   = "HFaD",
  "diethigh_fiber" = "HFiD"
)

qmp_labels <- c(
  "Microbial_Load" = "Fecal Bacterial Load\n [16S rRNA GCN/g per GI passage]",
  "Microbiota_Excretion" = "Bacterial Microbiota Excretion\n [16S rRNA GCN per GI passage]",
  "Transit_Time" = "GI Transit Time\n [min]",
  "Fecal_Mass" = "Fecal Mass\n [g per GI passage]"
)

plot_df_diet <- plot_df_diet %>%
  mutate(effect_color = ifelse(estimate < 0, "lower", "higher"))

plot_df_diet <- plot_df_diet %>%
  mutate(sig = FDR < 0.1)


plot_df_diet$variable <- factor(plot_df_diet$variable, levels = c("Transit_Time", "Fecal_Mass", "Microbial_Load", "Microbiota_Excretion"))
plot_df_diet$term <- factor(plot_df_diet$term, levels = c("diethigh_fiber", "diethigh_fat", "dietfasting"))


pdf("QMPs_LM_diet_norm.pdf", width = 16, height=6)
ggplot(plot_df_diet, aes(y = term, x = estimate, color = effect_color, shape = sig, fill = effect_color)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = lower.CL, xmax = upper.CL), height = 0.2, color = "grey40") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Effect size (estimate, 95% CI)",
    y = "",
    title = ""
  ) +
  scale_shape_manual(
    values = c("TRUE" = 21, "FALSE" = 1)) +
  facet_wrap(~ variable, ncol = 4, labeller = as_labeller(qmp_labels)) +
  scale_fill_manual(
    values = c("lower" = "red", "higher" = "royalblue")
  ) +
  scale_color_manual(
    values = c("lower" = "red", "higher" = "royalblue"),
    labels = c("lower" = "Lower than control", "higher" = "Higher than control"),
  ) +
  scale_y_discrete(labels = diet_labels) + 
   scale_x_continuous(limits = c(-0.75, 0.75),  breaks = c(-0.5, 0, 0.5))+
  Settings_forest
dev.off()

```


##2. TRF study
###2.1 Preparation
```{r}
vars_to_normalize <- c("Fecal_Mass", "Microbiota_Excretion", "Microbial_Load", "Transit_Time")

QMPs_ileum_TRF_fold <- TRF_data_ileum %>%
  group_by(project) %>%
  mutate(across(
    all_of(vars_to_normalize),
    ~ .x / mean(.x[diet == "control" & genotype_group == "WT"], na.rm = TRUE)
  )) %>%
  ungroup()

QMPs_colon_TRF_fold <- TRF_data_colon %>%
  group_by(project) %>%
  mutate(across(
    all_of(vars_to_normalize),
    ~ .x / mean(.x[diet == "control" & genotype_group == "WT"], na.rm = TRUE)
  )) %>%
  ungroup()

QMPs_ileum_TRF_fold <- QMPs_ileum_TRF_fold %>%
  mutate(diet_genotype = paste(diet, genotype_group, sep = "_"))
```

###2.2 Linear Regression
```{r}
variables <- c("Transit_Time", "Fecal_Mass", "Microbial_Load", "Microbiota_Excretion")

effect_results_TRF <- list()
emmeans_results_TRF <- list()

for (var in variables) {
  lm_data <- QMPs_ileum_TRF_fold %>%
    mutate(
      logY = log10(.data[[var]]),
      diet_genotype = factor(diet_genotype),                
      )
  fit <- lm(logY ~ diet_genotype, data = lm_data)
  
  res <- tidy(fit) %>%
    filter(term != "(Intercept)") %>%
    mutate(variable = var)
  effect_results_TRF[[var]] <- res
  
  emm <- emmeans(fit, ~ diet_genotype)
  
  pairs_res <- pairs(emm) %>% as.data.frame() %>% mutate(variable = var)
  emmeans_results_TRF[[var]] <- pairs_res
}

effect_results_TRF <- bind_rows(effect_results_TRF)
emmeans_results_TRF <- bind_rows(emmeans_results_TRF)

effect_results_TRF <- effect_results_TRF %>%
  group_by(variable) %>%
  mutate(FDR = p.adjust(p.value, method = "fdr")) %>%
  ungroup()
```

###2.3 Plot
```{r}
effect_results_TRF <- filter(effect_results_TRF, term != "diet_genotypeTRF_IL10-/-") # only plot effect of TRF on WT mice and genotype effect (--> exclude IL10KO mice with TRF)

plot_df_TRF <- effect_results_TRF %>%
  mutate(
    lower.CL = estimate - 1.96 * std.error,
    upper.CL = estimate + 1.96 * std.error
  )

TRF_labels <- c(
  "diet_genotypeTRF_WT"    = "TRF",
  "diet_genotypecontrol_IL10-/-"   = "Fasting"
                                      
)

qmp_labels <- c(
  "Microbial Load" = "Fecal Bacterial Load\n [16S rRNA GCN/g per GI passage]",
  "Microbiota Excretion" = "Bacterial Microbiota Excretion\n [16S rRNA GCN per GI passage]",
  "Transit Time" = "GI Transit Time\n [min]",
  "Fecal Mass" = "Fecal Mass\n [g per GI passage]"
)

plot_df_TRF <- plot_df_TRF %>%
  mutate(effect_color = ifelse(estimate < 0, "lower", "higher"))

plot_df_TRF <- plot_df_TRF %>%
  mutate(sig = FDR < 0.1)

plot_df_TRF$variable <- factor(plot_df_TRF$variable, levels = c("Transit_Time", "Fecal_Mass", "Microbial_Load", "Microbiota_Excretion"))


pdf("QMPs_LM_TRF_norm.pdf", width = 16, height=6)
ggplot(plot_df_TRF, aes(y = term, x = estimate, color = effect_color, shape = sig, fill = effect_color)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = lower.CL, xmax = upper.CL), height = 0.2, color = "grey40") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Effect size (estimate, 95% CI)",
    y = "",
    title = ""
  ) +
  scale_shape_manual(
    values = c("TRUE" = 21, "FALSE" = 1)) +
  facet_wrap(~ variable, ncol = 4, labeller = as_labeller(qmp_labels)) +
  scale_fill_manual(
    values = c("lower" = "red", "higher" = "royalblue")
  ) +
  scale_color_manual(
    values = c("lower" = "red", "higher" = "royalblue"),
    labels = c("lower" = "Lower than control", "higher" = "Higher than control"),
  ) +
  scale_y_discrete(labels = TRF_labels) + 
   scale_x_continuous(limits = c(-0.75, 0.75),  breaks = c(-0.5, 0, 0.5))+
  Settings_forest
dev.off()
```

# Figure 3B-D - QMP Linear regressions

## 1. Preparation
```{r}
meta_comb_passage <- metadata_QP_passage %>% filter(project == "diet_study" | project == "TRF_study") %>% filter(genotype_group == "WT") 
meta_comb_passage <- meta_comb_passage %>% filter(mouse != "529") # 63 samples 


# normalized fold change
qmp_vars <- c("Microbiota_Excretion", "Microbial_Load", "Transit_Time", "Fecal_Mass")

meta_comb_passage_norm <- meta_comb_passage %>%
  group_by(project) %>%
  mutate(across(all_of(qmp_vars),
                ~ .x / mean(.x[diet == "control"], na.rm = TRUE),
                .names = "{.col}_norm")) %>%
  ungroup()

table(meta_comb_passage_norm$diet)

meta_comb_passage_norm <- meta_comb_passage_norm %>%
  # for control diet of TRF study, rename into AL
  mutate(diet = case_when(diet == "control" & project == "TRF_study" ~ "AL", TRUE ~ diet)) # 

table(meta_comb_passage_norm$diet)

```


## 2. Stats


```{r}
# 1) define pairs to associate
pairs_to_run <- data.frame(
  Y = c("Microbiota_Excretion_norm", "Microbiota_Excretion_norm", "Microbiota_Excretion_norm", "Microbial_Load_norm","Microbial_Load_norm", "Fecal_Mass_norm"),
  X = c("Microbial_Load_norm", "Fecal_Mass_norm", "Transit_Time_norm", "Fecal_Mass_norm", "Transit_Time_norm", "Transit_Time_norm"),
  stringsAsFactors = FALSE)


# 2) fit models: controlled for diet (con) and unadjusted (unc), and extract values

results_raw <- list()

for (k in 1:nrow(pairs_to_run)) {
  
  y_var     <- pairs_to_run$Y[k]
  x_var     <- pairs_to_run$X[k]
  pair_name <- paste(y_var, x_var, sep = "_vs_")
  
  df_pair <- meta_comb_passage_norm[, c(y_var, x_var, "diet")]
  df_pair <- df_pair[complete.cases(df_pair), ]
  
  # log10 transform
  df_pair[[y_var]] <- log10(df_pair[[y_var]])
  df_pair[[x_var]] <- log10(df_pair[[x_var]])
  
  # remove infinite values
  df_pair <- df_pair[is.finite(df_pair[[y_var]]) & 
                     is.finite(df_pair[[x_var]]), ]
  
  # fit models
  lm_unc <- lm(as.formula(paste(y_var, "~", x_var)),           data = df_pair)
  lm_con <- lm(as.formula(paste(y_var, "~", x_var, "+ diet")), data = df_pair)
  
  # --- uncontrolled stats ---
  s_unc    <- summary(lm_unc)
  est_unc  <- s_unc$coefficients[x_var, "Estimate"]
  se_unc   <- s_unc$coefficients[x_var, "Std. Error"]
  std_beta_unc <- standardize_parameters(lm_unc)$Std_Coefficient[2]
  r2_unc   <- s_unc$r.squared
  p_unc    <- s_unc$coefficients[x_var, "Pr(>|t|)"]
  
  # --- controlled stats ---
  s_con    <- summary(lm_con)
  est_con  <- s_con$coefficients[x_var, "Estimate"]
  se_con   <- s_con$coefficients[x_var, "Std. Error"]
  std_beta_con <- standardize_parameters(lm_con)$Std_Coefficient[2]
  r2_con   <- s_con$r.squared
  p_con    <- s_con$coefficients[x_var, "Pr(>|t|)"]
  
  results_raw[[pair_name]] <- data.frame(
    Y           = y_var,
    X           = x_var,
    Est_unc     = est_unc,
    SE_unc      = se_unc,
    StdBeta_unc = std_beta_unc,
    R2_unc      = r2_unc,
    P_unc       = p_unc,
    Est_con     = est_con,
    SE_con      = se_con,
    StdBeta_con = std_beta_con,
    R2_con      = r2_con,
    P_con       = p_con,
    stringsAsFactors = FALSE)
}


# 3) combine model output, comüute CIs and apply BH correction

lm_table <- bind_rows(results_raw) %>%
  dplyr::mutate(
    lower.CL_unc = Est_unc - 1.96 * SE_unc,
    upper.CL_unc = Est_unc + 1.96 * SE_unc,
    lower.CL_con = Est_con - 1.96 * SE_con,
    upper.CL_con = Est_con + 1.96 * SE_con) %>%
  
   dplyr::mutate(
    P_adj_unc = p.adjust(P_unc, method = "BH"),
    P_adj_con = p.adjust(P_con, method = "BH")) %>%
  
  add_significance("P_adj_unc") %>%
  add_significance("P_adj_con") %>%
  
   dplyr::mutate(
     across(c(Est_unc, SE_unc, StdBeta_unc, R2_unc, lower.CL_unc, upper.CL_unc, 
              Est_con, SE_con, StdBeta_con, R2_con, lower.CL_con, upper.CL_con), ~ round(.x, 3)),
     across(c(P_unc, P_adj_unc, P_con, P_adj_con), ~ signif(.x, 3))) %>%
  
   dplyr::select(Y, X,
         Est_unc, SE_unc, StdBeta_unc, lower.CL_unc, upper.CL_unc,
         R2_unc, P_unc, P_adj_unc, P_adj_unc.signif,
         Est_con, SE_con, StdBeta_con, lower.CL_con, upper.CL_con,
         R2_con, P_con, P_adj_con, P_adj_con.signif)

# pivot longer
lm_table_long <- lm_table %>%
  rename(Sig_unc = P_adj_unc.signif, Sig_con = P_adj_con.signif) %>%
  pivot_longer(
    cols = c(Est_unc, SE_unc, StdBeta_unc, lower.CL_unc, upper.CL_unc, R2_unc, P_unc, P_adj_unc, Sig_unc,
             Est_con, SE_con, StdBeta_con, lower.CL_con, upper.CL_con, R2_con, P_con, P_adj_con, Sig_con),
    names_to  = c(".value", "Model"),
    names_pattern = "(.*)_(unc|con)") %>%
  mutate(Model = ifelse(Model == "unc", "Uncontrolled", "Controlled")) %>%
  arrange(Y, X, Model)

```

Notes to the model:  
- extracted p values refer to the X variables only, so only the QMPs - but not diet!  
- extracted std. coefficients are also for the X variable only 
- extracted R2 refers to the whole model, so for the controlled model it includes the contribution of diet as well


## 3. Plots

Add trendlines across complete dataset, but only for associations that remain significant after controlling for diet 

```{r}
Diet_AL_TRF_colors <- c(
  "control"     = "#85C1E9",
  "fasting"     = "#AF7AC5",
  "high_fat"    = "#F39C12",
  "high_fiber"  = "#52BE80",
  "AL"          = "#3498DB",
  "TRF"         = "#2471A3")

Diet_AL_TRF_labels <- c(
  "control"     = "RCD (diet study)",
  "fasting"     = "Fasting",
  "high_fat"    = "HFaD",
  "high_fiber"  = "HFiD",
  "AL"          = "RCD-AL (TRF study)",
  "TRF"         = "RCD-TRF")
```


```{r}
# microbiota excretion (GCN) - fecal microbial load (GCN/g) => diet-independent
p1 <- meta_comb_passage_norm %>% 
  
  ggplot(aes(Microbial_Load_norm, Microbiota_Excretion_norm)) +
  geom_point(shape = 16, size = 4, aes(fill = diet, color = diet)) + 
 
  geom_smooth(method = 'lm', color = 'black', linewidth = 1, se = F) +
  
  scale_color_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  scale_fill_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  
  labs(x = paste0("<span style='font-size: 25pt'>Microbial Load</span><br><span style='font-size: 20pt'>[16S rRNA GCN/g per GI Passage]</span>"), 
       y = paste0("<span style='font-size: 25pt'>Microbiota Excretion</span><br><span style='font-size: 20pt'>[16S rRNA GCN per GI Passage]</span>")) +
  Settings_corr +  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown(), legend.position = "none")

# microbiota excretion (GCN) - fecal mass => diet-masked
p2 <- meta_comb_passage_norm %>% 
  
  ggplot(aes(Fecal_Mass_norm, Microbiota_Excretion_norm)) +
  geom_point(shape = 16, size = 4, aes(fill = diet, color = diet)) + 
  
  #geom_smooth(method = 'lm', color = 'grey50', linetype = "dashed", linewidth = 1, se = F) +
  
  scale_color_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  scale_fill_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  
  labs(x = paste0("<span style='font-size: 25pt'>Fecal Mass</span><br><span style='font-size: 20pt'>[g per GI Passage]</span>"), 
       y = paste0("<span style='font-size: 25pt'>Microbiota Excretion</span><br><span style='font-size: 20pt'>[16S rRNA GCN per GI Passage]</span>")) +
  Settings_corr +  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown(), legend.position = "none")

# fecal microbial load (GCN/g) - fecal mass => diet-driven 
p3 <- meta_comb_passage_norm %>% 
  
  ggplot(aes(Fecal_Mass_norm, Microbial_Load_norm)) +
  geom_point(shape = 16, size = 4, aes(fill = diet, color = diet)) + 
  
  #geom_smooth(method = 'lm', color = 'grey50', linetype = "dashed", linewidth = 1, se = F) +
  
  scale_color_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  scale_fill_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +

  labs(x = paste0("<span style='font-size: 25pt'>Fecal Mass</span><br><span style='font-size: 20pt'>[g per GI Passage]</span>"), 
       y = paste0("<span style='font-size: 25pt'>Microbial Load</span><br><span style='font-size: 20pt'>[16S rRNA GCN/g per GI Passage]</span>")) +
  Settings_corr +  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown(), legend.position = "none")


```

```{r}
# transit - microbiota excretion => diet-independent
p4 <- meta_comb_passage_norm %>% 
  
  ggplot(aes(Transit_Time_norm, Microbiota_Excretion_norm)) +
  geom_point(shape = 16, size = 4, aes(fill = diet, color = diet)) + 
  
  geom_smooth(method = 'lm', color = 'black', linewidth = 1, se = F) +
  
  scale_color_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  scale_fill_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +

  labs(x = paste0("<span style='font-size: 25pt'>GI Transit Time</span><br><span style='font-size: 20pt'>[min]</span>"), 
       y = paste0("<span style='font-size: 25pt'>Microbiota Excretion</span><br><span style='font-size: 20pt'>[16S rRNA GCN per GI Passage]</span>")) +
  Settings_corr +  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown(), legend.position = "none")

# transit - microbial load (GCN/g) => diet-independent
p5 <- meta_comb_passage_norm %>%
  
  ggplot(aes(Transit_Time_norm, Microbial_Load_norm)) +
  geom_point(shape = 16, size = 4, aes(fill = diet, color = diet)) + 
  
  geom_smooth(method = 'lm', color = 'black', linewidth = 1, se = F) +
  
  scale_color_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  scale_fill_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +

  labs(x = paste0("<span style='font-size: 25pt'>GI Transit Time</span><br><span style='font-size: 20pt'>[min]</span>"), 
       y = paste0("<span style='font-size: 25pt'>Microbial Load</span><br><span style='font-size: 20pt'>[16S rRNA GCN/g per GI Passage]</span>")) +
  Settings_corr +  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown(), legend.position = "none")

# transit - fecal mass => no association 
p6 <- meta_comb_passage_norm %>% 

  ggplot(aes(Transit_Time_norm, Fecal_Mass_norm)) +
  geom_point(shape = 16, size = 4, aes(fill = diet, color = diet)) + 
  
  scale_color_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +
  scale_fill_manual(values = Diet_AL_TRF_colors,  name = "Diet", labels = Diet_AL_TRF_labels) +

  labs(x = paste0("<span style='font-size: 25pt'>GI Transit Time</span><br><span style='font-size: 20pt'>[min]</span>"), 
       y = paste0("<span style='font-size: 25pt'>Fecal Mass</span><br><span style='font-size: 20pt'>[g per GI Passage]</span>")) +
  Settings_corr +  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown(), legend.position = "none")

```


```{r}
# export plot
pdf("QMP_QMP_corr_grid_norm.pdf", width = 20, height = 18)
cowplot::plot_grid(plotlist = cowplot::align_plots(p1, p2, p3, p4, p5, p6, align = "hv", axis = "tblr"), ncol = 3, rel_widths = c(1, 1, 1))
dev.off()

```
