---
title: "Figure 6"
output: html_document
date: "2026-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figure 6A - PCoA Plot diet & TRF study 

## 1. Prepare data
```{r}
ASV_QP_interval <- rownames_to_column(ASV_count_combined, var = "sampleID")

ASV_meta_interval <- right_join(ASV_QP_interval, metadata_QP_interval) %>%
  tidyr::drop_na()

ASV_meta_interval <- filter(ASV_meta_interval, (project == "diet_study") |  (project == "TRF_study"))

ASV_interval_for_mean <- ASV_meta_interval[, -c(11295:11311)]
```

```{r}
ASV_mean <- ASV_interval_for_mean %>%
  group_by(mouse) %>%    
  summarise(across(-sampleID, mean))

ASV_mean <- ASV_mean %>%
  column_to_rownames("mouse")
```

```{r}
metadata_PCoA <- metadata_QP_passage %>%
  mutate(diet_project = paste(diet, project, sep = "_"))

metadata_PCoA <- metadata_PCoA %>%
  mutate(diet_project_genotype = paste(diet_project, genotype_group, sep = "_"))

#metadata_diet_WP3 <- metadata_diet_WP3 %>%
#  rownames_to_column("SampleID")
```

## 2. Calculate Bray-Curtis Dissimilarity
```{r}
ASV_mean <- as.matrix(ASV_mean)

ASV_relab<- sweep(ASV_mean, 1, rowSums(ASV_mean),'/')
dim(ASV_relab)

ASV_relab <- as.data.frame(ASV_relab)

ASV_relab_filt <- ASV_relab %>%
  rownames_to_column("mouse") %>%
  filter(mouse %in% metadata_PCoA$mouse) %>%
  column_to_rownames("mouse")

## calculate Bray-Curtis Dissimilarity
d.bray <-vegdist(ASV_relab_filt, na.rm = T, method = "bray")
matrix.bray_ANOSIM <- as.matrix(d.bray)
```

## 3. Visualization
```{r}
# calculate PCoA values based on Bray-Curtis Dissimilarity
pcoa.bray <- cmdscale(d.bray, k=2, eig = TRUE)

# Generate percentages % for plot axis
explainvar1 <- round(pcoa.bray$eig[1]/sum(pcoa.bray$eig), 3)*100
explainvar2 <- round(pcoa.bray$eig[2]/sum(pcoa.bray$eig), 3)*100

pcoa_map <- merge(metadata_PCoA, pcoa.bray$points, by.x = "mouse", by.y = 0)


pcoa_map$diet_project_genotype <- factor(pcoa_map$diet_project_genotype, levels = c("control_diet_study_WT", "fasting_diet_study_WT", "high_fat_diet_study_WT", "high_fiber_diet_study_WT", "control_TRF_study_WT", "TRF_TRF_study_WT", "control_TRF_study_IL10-/-", "TRF_TRF_study_IL10-/-"))

pdf("PCoA_both_projects_mean.pdf", width = 8, height=6)
ggplot(pcoa_map, aes(V1, V2, colour= diet_project_genotype, shape = project))+
  geom_point(size=4)+
  xlab(paste("PCoA 1 (" ,explainvar1, "%)", sep = "")) +
  ylab(paste("PCoA 2 (" ,explainvar2, "%)", sep = "")) +
  scale_fill_manual(
    values = c("#85C1E9", "#AF7AC5", "#F39C12", "#52BE80", "#3498DB", "#2471A3", "#E74C3C", "#A93226"),
    labels = c("control_diet_study_WT" = "Control", "fasting_diet_study_WT" = "Fasting", "high_fat_diet_study_WT" = "High Fat", "high_fiber_diet_study_WT" = "High Fiber", "control_TRF_study_WT" = "WT ad libitum", "TRF_TRF_study_WT" = "WT TRF", "control_TRF_study_IL10-/-" = expression("IL-10"^"-/-"~"ad libitum"), "TRF_TRF_study_IL10-/-" = expression("IL-10"^"-/-"~"TRF")),
                     name = "") +
  scale_color_manual(
    values = c("#85C1E9", "#AF7AC5", "#F39C12", "#52BE80", "#3498DB", "#2471A3", "#E74C3C", "#A93226"),
      labels = c("control_diet_study_WT" = "Control", "fasting_diet_study_WT" = "Fasting", "high_fat_diet_study_WT" = "High Fat", "high_fiber_diet_study_WT" = "High Fiber", "control_TRF_study_WT" = "WT ad libitum", "TRF_TRF_study_WT" = "WT TRF", "control_TRF_study_IL10-/-" = expression("IL-10"^"-/-"~"ad libitum"), "TRF_TRF_study_IL10-/-" = expression("IL-10"^"-/-"~"TRF")),
                     name = "") +
  scale_shape_manual(
    values = c(16, 17), 
    labels = c("TRF_study" = "TRF study", "diet_study" = "diet study"),
    name = ""
  ) +
  Settings_PCoA 
dev.off()

```

##4.  Statistical testing
diet
```{r}
order_anosim <- rownames(matrix.bray_ANOSIM)
metadata_PCoA <- metadata_PCoA[match(order_anosim, metadata_PCoA$mouse), ]

## perform ANOSIM 
anosim(matrix.bray_ANOSIM, grouping = metadata_PCoA$diet, permutations = 999)
```

project
```{r}
order_anosim <- rownames(matrix.bray_ANOSIM)
metadata_PCoA <- metadata_PCoA[match(order_anosim, metadata_PCoA$mouse), ]

## perform ANOSIM 
anosim(matrix.bray_ANOSIM, grouping = metadata_PCoA$project, permutations = 999)
```

# Figure 6B - dbRDA diet & TRF study 

## 1. Preparation

Load and prepare data 
```{r}

meta_L6_dbRDA <- meta_mean_L6_relAb_combined %>% 
  filter((project == "diet_study" | project == "TRF_study")) %>% 
  dplyr::select(-mouse, -passage_cycle, -ZT, -clock_group, -disrupted_clock) %>%  
  
  # genotype variables
  mutate(genotype_all = case_when(
    genotype == "black6" ~ "WT_Diet",
    (genotype == "BALBc" & genotype_group == "WT") ~ "WT_TRF",
    genotype_group == "IL10-/-" ~ "IL10-/-",
    TRUE ~ NA_character_)) %>% 
  mutate(genotype_2 = genotype_group) %>% 
  
  # diet and feeding type variables 
  mutate(diet_type = case_when(
    (diet == "control" | diet == "TRF" | diet == "fasting") ~ "regular_chow", # TRF and fasting also need to be classified as regular chow
    diet == "high_fat" ~ "high_fat",
    diet == "high_fiber" ~ "high_fiber",
    TRUE ~ NA_character_)) %>% 
  mutate(feeding_type = case_when(
    (diet == "control" | diet == "high_fat" | diet == "high_fiber") ~ "ad_libitum", # same here for high fat and high fiber, need to be classified as ad libitum
    diet == "TRF" ~ "TRF", 
    diet == "fasting" ~"fasting", 
    TRUE ~ NA_character_)) %>% 
  
  # de-select 
  dplyr::select(-genotype_group, -genotype, -diet, -colitis) %>% 
  relocate(c(genotype_all, genotype_2, diet_type, feeding_type), .after = "project") 

# in this set-up, the variable feeding_type also contains the high fat and high fiber diets as controls since they are also ad libitum -> universally applicable variables

```


Calculate Bray Curtis Dissimilarity
```{r}
# subset (make sure sampleIDs are rownames)
meta_L6_dbRDA <- meta_L6_dbRDA %>% column_to_rownames("sampleID")
meta_dbRDA <- meta_L6_dbRDA[1:11] # metadata columns 
L6_relAb_dbRDA <- meta_L6_dbRDA[12:177] # taxa columns 

# factor levels
meta_dbRDA[1:7] <- lapply(meta_dbRDA[1:7], as.factor)
# log10 transform QMPs
meta_dbRDA[8:11] <- log10(meta_dbRDA[8:11])

# BC
L6_bray_dbRDA <- vegdist(L6_relAb_dbRDA, method = "bray", na.rm = T) # matrix 97x97
L6_bray_mx_dbRDA <- as.matrix(L6_bray_dbRDA)
```


## 2. dbRDA model

assess explanatory of each possible variable separately 

```{r}
# Initialize results & model storage
ind_dbRDA_all <- data.frame()
models <- list()

# Loop through metadata variables
for (ii in 1:ncol(meta_dbRDA)) {  
  dbrda <- dbrda(L6_bray_dbRDA ~ meta_dbRDA[, ii], distance = "bray", na.action = na.omit)  # Fit each metadata variable to the BC matrix
  an <- anova.cca(dbrda, permutations = 9999)  # ANOVA on constrained ordination
  pval <- an["Pr(>F)"][[1]][[1]]  # Extract p-value
  R2adj <- RsquareAdj(dbrda)[[2]]  # Extract adjusted R²
  ind_dbRDA_all <- rbind(ind_dbRDA_all, cbind(R2adj, pval))  # Store results
  models[[colnames(meta_dbRDA)[ii]]] <- dbrda # Save models 
}

# but add variable names & significance levels
ind_dbRDA_all <- ind_dbRDA_all %>% mutate(variable = colnames(meta_dbRDA)) %>% relocate(variable, .before = R2adj) %>% 
  mutate(padj = p.adjust(pval, method = "BH")) %>% add_significance("padj")   

# add category for plotting with ifelse based on variable_colors below
ind_dbRDA_all <- ind_dbRDA_all %>% 
  mutate(category = case_when(
    variable %in% c("diet_type", "feeding_type") ~ "Diet/Feeding",
    variable %in% c("genotype_all", "genotype_2") ~ "Genotype",
    variable %in% c("Fecal_Mass", "Microbial_Load", "Microbiota_Excretion", "Transit_Time") ~ "QMP",
    variable %in% c("cage", "litter", "project") ~ "Experimental",
    TRUE ~ "other"
  ))

# order for plot
ind_dbRDA_all <- ind_dbRDA_all[order(ind_dbRDA_all$R2adj, decreasing = F),]
ind_dbRDA_all$variable <- factor(ind_dbRDA_all$variable, levels = ind_dbRDA_all$variable)
print(ind_dbRDA_all) 

```


## 3. Plot

Barplot
```{r}

variable_colors <- c("Diet/Feeding" = "#7690a0", "Experimental" = "#92ab93", "Genotype" = "#e5e1d9", "QMP" = "#a38e8e")

# plot with custom colors
pdf("dbRDA_barplot_all_variables.pdf", width = 18, height = 10)

ggplot(ind_dbRDA_all, aes(x = variable, y = R2adj, fill = category)) +
  geom_bar(stat = "identity", position = position_dodge2(), alpha = 0.8) +
  geom_text(aes(label = padj.signif), position = position_dodge(width = .9), hjust = -0.5, size = 8) +
  scale_fill_manual(values = variable_colors, name = "Categories") + 
  scale_x_discrete(labels = c("cage" = "Cage", "litter" = "Litter", "genotype_all" = expression("Mouse Strain"), "project" = "Mouse Study", "feeding_type" = "Feeding Type", "Fecal_Mass" = "Fecal Mass", 
                              "diet_type" = "Diet Type", "Microbial_Load" = "Microbial Load", "Transit_Time" = "GI Transit Time", "genotype_2" = expression("Genotype (WT, IL-10"^"-/-" * ")"), 
                              "Microbiota_Excretion" = "Microbiota Excretion")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  coord_cartesian(ylim = c(0, 1)) +
  Settings_barplot + theme(plot.title = element_text(face = "bold", size = 25), legend.position = "right") +
  labs(title = "Individual Effects on Relative Microbiota Composition", y = expression("Explained Variance (Adj. R"^"2" * ")"), x = "Variables") +
  coord_flip()  # Flip for better readability

dev.off()

```



# Figure 6C - dbRDA diet & TRF study controlled for project

## 1. Preparation

as in Figure 6B, 1. Preparation

## 2. dbRDA model
```{r}
# Initialize output
ind_dbRDA_all_project <- data.frame()
models <- list()

# All variables EXCEPT project

vars_all <- setdiff(colnames(meta_dbRDA), "project")

# Loop
for (var in vars_all) {
  form <- as.formula(paste0("L6_bray_dbRDA ~ ", var, " + Condition(project)"))
  dbrda <- dbrda(form, data = meta_dbRDA, distance = "bray", na.action = na.omit)
  an <- anova.cca(dbrda, permutations = 9999)
  pval <- an["Pr(>F)"][[1]][[1]]
  R2adj <- RsquareAdj(dbrda)[[2]]
  
  ind_dbRDA_all_project <- rbind(ind_dbRDA_all_project, cbind(R2adj, pval))
  models[[var]] <- dbrda
}

# Add annotations
ind_dbRDA_all_project <- ind_dbRDA_all_project %>%
  mutate(variable = vars_all) %>%
  relocate(variable, .before = R2adj) %>%
  mutate(padj = p.adjust(pval, method = "BH")) %>%
  add_significance("padj")


# add category for plotting with ifelse based on variable_colors below
ind_dbRDA_all_project <- ind_dbRDA_all_project %>% 
  mutate(category = case_when(
    variable %in% c("diet_type", "feeding_type") ~ "Diet/Feeding",
    variable %in% c("genotype_all", "genotype_2") ~ "Genotype",
    variable %in% c("Fecal_Mass", "Microbial_Load", "Microbiota_Excretion", "Transit_Time") ~ "QMP",
    variable %in% c("cage", "litter") ~ "Experimental",
    TRUE ~ "other"
  ))

# order for plot
ind_dbRDA_all_project <- ind_dbRDA_all_project[order(ind_dbRDA_all_project$R2adj, decreasing = F),]
ind_dbRDA_all_project$variable <- factor(ind_dbRDA_all_project$variable, levels = ind_dbRDA_all_project$variable)
print(ind_dbRDA_all_project)

```


## 3. Plots

Barplot - all variables, controlled for project = *Suppl. Figure XX*
```{r}
variable_colors <- c("Diet/Feeding" = "#7690a0", "Experimental" = "#92ab93", "Genotype" = "#e5e1d9", "QMP" = "#a38e8e")

pdf("dbRDA_barplot_all_variables_project.pdf", width = 18, height = 10)

ggplot(ind_dbRDA_all_new_project, aes(x = variable, y = R2adj, fill = category)) +
  geom_bar(stat = "identity", position = position_dodge2(), alpha = 0.8) +
  geom_text(aes(label = padj.signif), position = position_dodge(width = .9), hjust = -0.5, size = 8) +
  scale_fill_manual(values = variable_colors, name = "Categories") + 
  scale_x_discrete(labels = c("cage" = "Cage", "litter" = "Litter", "genotype_all" = expression("Mouse Strain"), "project" = "Mouse Study", "feeding_type" = "Feeding Type", "stool_cycle" = "Fecal Mass", 
                              "diet_type" = "Diet Type", "density_new" = "Microbial Load", "passage_transit" = "GI Transit Time", "genotype_2" = expression("Genotype (WT, IL-10"^"-/-" * ")"), 
                              "GCN_cycle" = "Microbiota Excretion")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  coord_cartesian(ylim = c(0, 1)) +
  Settings_barplot + theme(plot.title = element_text(face = "bold", size = 25), plot.subtitle = element_text(size = 22, hjust = 0.5), legend.position = "right") +
  labs(title = "Individual Effects on Relative Microbiota Composition", subtitle = "Controlled for Mouse Study", y = expression("Explained Variance (Adj. R"^"2" * ")"), x = "Variables") +
  coord_flip()  # Flip for better readability

dev.off()
```


Barplot - subset plot with only QMP variables 
```{r}
variable_colors <- c("Diet/Feeding" = "#7690a0", "Experimental" = "#92ab93", "Genotype" = "#e5e1d9", "QMP" = "#a38e8e")


pdf("dbRDA_barplot_QMP_project.pdf", width = 10, height = 6)

ggplot(ind_dbRDA_all_new_project %>% filter(category == "QMP"), aes(x = variable, y = R2adj, fill = category)) +
  geom_bar(stat = "identity", position = position_dodge2(), alpha = 0.8) +
  geom_text(aes(label = padj.signif), position = position_dodge(width = .9), hjust = -0.5, size = 8) +
  scale_fill_manual(values = variable_colors, name = "Categories") + 
    scale_x_discrete(labels = c("stool_cycle" = "Fecal Mass", "density_new" = "Microbial Load", "passage_transit" = "GI Transit Time", "GCN_cycle" = "Microbiota Excretion")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  coord_cartesian(ylim = c(0, 1)) +
  Settings_barplot + theme(plot.title = element_text(face = "bold", size = 25), plot.subtitle = element_text(size = 22, hjust = 0.5), legend.position = "none") +
  labs(title = "Individual Effects on Relative Microbiota Composition", subtitle = "QMP variables only, controlled for Mouse Study", y = expression("Explained Variance (Adj. R"^"2" * ")"), x = "Variables") +
  coord_flip()  # Flip for better readability

dev.off()

```



# Figure 6D - QMP - Taxa Correlation

##1. Preparation

```{r}
L6_diet_trf <- filter(meta_mean_L6_relAb_combined, project == "diet_study" | project == "TRF_study" & genotype_group == "WT")
```

```{r}
vars_to_normalize <- c(
  "Fecal_Mass",
  "Microbiota_Excretion",
  "Microbial_Load",
  "Transit_Time",
  grep("^g_|^f_|^o_|^c_|^p_|^k_", names(L6_diet_trf), value = TRUE)
)

QMPs_taxa_fold <- L6_diet_trf %>%
  group_by(project) %>%
  mutate(across(
    all_of(vars_to_normalize),
    ~ .x / mean(.x[diet == "control"], na.rm = TRUE)
  )) %>%
  ungroup()

QMPs_taxa_fold <- QMPs_taxa_fold %>%
  select(where(~ !all(is.na(.))))
```

##2. Corrleations

### 2.1  Microbiota Excretion [16S rRNA GCN per GI Passage]
Stats
```{r}
L6_corr_excr_filt_WT <- data.frame()

for (i in taxa_to_keep) {

    tryCatch({

        non_zero_data <- QMPs_taxa_fold[QMPs_taxa_fold[[i]] > 0 & !is.na(QMPs_taxa_fold[[i]]), ]
        num_points <- nrow(non_zero_data)

        if (num_points >= 10) {

            x <- non_zero_data[[i]]
            y <- non_zero_data$Microbiota_Excretion

            tmp_corr_spearman <- cor.test(x, y, method = "spearman", exact = FALSE)

            L6_corr_excr_filt_WT <- rbind(
                L6_corr_excr_filt_WT,
                data.frame(
                    taxa = i,
                    Rho = tmp_corr_spearman$estimate,
                    P_value = tmp_corr_spearman$p.value,
                    n_points = num_points
                )
            )

        } else {

            L6_corr_excr_filt_WT <- rbind(
                L6_corr_excr_filt_WT,
                data.frame(
                    taxa = i,
                    Rho = NA,
                    P_value = NA,
                    n_points = num_points
                )
            )
        }

    }, error = function(e) return(NA))

}


L6_corr_excr_filt_WT$P_adjusted <- p.adjust(L6_corr_excr_filt_WT$P_value, method = "BH")

L6_corr_excr_filt_WT <- L6_corr_excr_filt_WT %>%
  mutate(Sign. = case_when(
    P_adjusted >= 0.05 ~ "ns",
    P_adjusted < 0.05 & P_adjusted >= 0.01 ~ "*",
    P_adjusted < 0.01 & P_adjusted >= 0.001 ~ "**",
    P_adjusted < 0.001 & P_adjusted >= 0.0001 ~ "***",
    P_adjusted < 0.0001 ~ "****"
  ))

L6_corr_excr_filt_WT[,"Quantitative_Parameter"] <- "Microbiota Excretion [16s GCN/GI Passage]"
L6_corr_excr_filt_WT <- L6_corr_excr_filt_WT %>% relocate("Quantitative_Parameter") %>% relocate("n_points", .after = taxa)
row.names(L6_corr_excr_filt_WT) <- NULL


```

### 2.2 Microbial Load [16S rRNA GCN/ g stool]
Stats
```{r}
L6_corr_load_filt_WT <- data.frame()

for (i in unique(taxa_to_keep)) {

    tryCatch({
        non_zero_data <- QMPs_taxa_fold[QMPs_taxa_fold[[i]] > 0 & !is.na(QMPs_taxa_fold[[i]]), ]
        num_points <- nrow(non_zero_data)

        if (num_points >= 10) {

            x <- non_zero_data[[i]]
            y <- non_zero_data$Microbial_Load

            tmp_corr_spearman <- cor.test(x, y, method = "spearman", exact = FALSE)

            L6_corr_load_filt_WT <- rbind(
                L6_corr_load_filt_WT,
                data.frame(
                    taxa = i,
                    Rho = tmp_corr_spearman$estimate,
                    P_value = tmp_corr_spearman$p.value,
                    n_points = num_points
                )
            )

        } else {

            L6_corr_load_filt_WT <- rbind(
                L6_corr_load_filt_WT,
                data.frame(
                    taxa = i,
                    Rho = NA,
                    P_value = NA,
                    n_points = num_points
                )
            )
        }

    }, error = function(e) return(NA))

}

L6_corr_load_filt_WT$P_adjusted <- p.adjust(L6_corr_load_filt_WT$P_value, method = "BH")

L6_corr_load_filt_WT <- L6_corr_load_filt_WT %>%
  mutate(Sign. = case_when(
    P_adjusted >= 0.05 ~ "ns",
    P_adjusted < 0.05 & P_adjusted >= 0.01 ~ "*",
    P_adjusted < 0.01 & P_adjusted >= 0.001 ~ "**",
    P_adjusted < 0.001 & P_adjusted >= 0.0001 ~ "***",
    P_adjusted < 0.0001 ~ "****"
  ))

L6_corr_load_filt_WT$Quantitative_Parameter <- "Microbial Load [16s GCN/g stool]"

L6_corr_load_filt_WT <- L6_corr_load_filt_WT %>%
  relocate(Quantitative_Parameter) %>%
  relocate(n_points, .after = taxa)

row.names(L6_corr_load_filt_WT) <- NULL


```

### 2.3. Fecal Mass (g)
Stats
```{r}
L6_corr_stool_filt_WT <- data.frame()

for (i in unique(taxa_to_keep)) {
    tryCatch({

        # Remove NA and 0
        non_zero_data <- QMPs_taxa_fold[QMPs_taxa_fold[[i]] > 0 & !is.na(QMPs_taxa_fold[[i]]), ]
        num_points <- nrow(non_zero_data)
        
        if (num_points >= 10) {

            x <- non_zero_data[[i]]
            y <- non_zero_data$Fecal_Mass
            
            tmp_corr_spearman <- cor.test(x, y, method = "spearman", exact = FALSE)
            
            L6_corr_stool_filt_WT <- rbind(
                L6_corr_stool_filt_WT, 
                data.frame(
                    taxa = i,
                    Rho = tmp_corr_spearman$estimate,
                    P_value = tmp_corr_spearman$p.value,
                    n_points = num_points
                )
            )

        } else {

            L6_corr_stool_filt_WT <- rbind(
                L6_corr_stool_filt_WT, 
                data.frame(
                    taxa = i,
                    Rho = NA,
                    P_value = NA,
                    n_points = num_points
                )
            )
        }

    }, error = function(e) return(NA))
}

# FDR correction
L6_corr_stool_filt_WT$P_adjusted <- p.adjust(L6_corr_stool_filt_WT$P_value, method = "BH")

# Significance labels
L6_corr_stool_filt_WT <- L6_corr_stool_filt_WT %>%
  mutate(Sign. = case_when(
    P_adjusted >= 0.05 ~ "ns",
    P_adjusted < 0.05 & P_adjusted >= 0.01 ~ "*",
    P_adjusted < 0.01 & P_adjusted >= 0.001 ~ "**",
    P_adjusted < 0.001 & P_adjusted >= 0.0001 ~ "***",
    P_adjusted < 0.0001 ~ "****"
  ))

# Add parameter label and reorder columns
L6_corr_stool_filt_WT$Quantitative_Parameter <- "Fecal Mass [g]"

L6_corr_stool_filt_WT <- L6_corr_stool_filt_WT %>%
  relocate(Quantitative_Parameter) %>%
  relocate(n_points, .after = taxa)

row.names(L6_corr_stool_filt_WT) <- NULL


```

### 2.4. GI Transit Time (min)
Stats
```{r}
L6_corr_time_filt_WT <- data.frame()

for (i in unique(taxa_to_keep)) {
    tryCatch({
        
        non_zero_data <- QMPs_taxa_fold[QMPs_taxa_fold[[i]] > 0 & !is.na(QMPs_taxa_fold[[i]]), ]
        num_points <- nrow(non_zero_data)
        
        if (num_points >= 10) {  
            
            x <- non_zero_data[[i]]
            y <- non_zero_data$Transit_Time
            
            tmp_corr_spearman <- cor.test(x, y, method = "spearman", exact = FALSE)
            
            L6_corr_time_filt_WT <- rbind(
                L6_corr_time_filt_WT, 
                data.frame(
                    taxa = i,
                    Rho = tmp_corr_spearman$estimate,
                    P_value = tmp_corr_spearman$p.value,
                    n_points = num_points
                )
            )
            
        } else {
            
            L6_corr_time_filt_WT <- rbind(
                L6_corr_time_filt_WT, 
                data.frame(
                    taxa = i,
                    Rho = NA,
                    P_value = NA,
                    n_points = num_points
                )
            )
        }
        
    }, error = function(e) return(NA))
}
L6_corr_time_filt_WT$P_adjusted <- p.adjust(L6_corr_time_filt_WT$P_value, method = "BH")

L6_corr_time_filt_WT <- L6_corr_time_filt_WT %>% mutate(
    Sign. = case_when(
        P_adjusted >= 0.05 ~ "ns",
        P_adjusted < 0.05 & P_adjusted >= 0.01 ~ "*",
        P_adjusted < 0.01 & P_adjusted >= 0.001 ~ "**",
        P_adjusted < 0.001 & P_adjusted >= 0.0001 ~ "***",
        P_adjusted < 0.0001 ~ "****"
    )
)

L6_corr_time_filt_WT$Quantitative_Parameter <- "GI Transit Time [min]"

L6_corr_time_filt_WT <- L6_corr_time_filt_WT %>%
    relocate(Quantitative_Parameter) %>%
    relocate(n_points, .after = taxa)

row.names(L6_corr_time_filt_WT) <- NULL
```


## 3. Overview Plot Coefficients

```{r}
combined_df_WT <- bind_rows(
  L6_corr_time_filt_WT,
  L6_corr_stool_filt_WT,
  L6_corr_excr_filt_WT,
  L6_corr_load_filt_WT
) %>%
  mutate(QMP = case_when(
    Quantitative_Parameter == "GI Transit Time [min]" ~ "Transit_Time",
    Quantitative_Parameter == "Fecal Mass [g]" ~ "Fecal_Mass",
    Quantitative_Parameter == "Microbiota Excretion [16s GCN/GI Passage]" ~ "Microbiota_Excretion",
    Quantitative_Parameter == "Microbial Load [16s GCN/g stool]" ~ "Microbial_Load"
  ))


combined_df_WT$Rho <- as.numeric(combined_df_WT$Rho)

combined_df_WT$Significance <- ifelse(combined_df_WT$Sign. != "ns", TRUE, FALSE)

combined_df_WT <- combined_df_WT %>%
  arrange(taxa)
```

safe corr. results
```{r}
write.table(combined_df_WT, "Rel_abundanca_taxa_qmps_correlations_fold.txt", 
            row.names = FALSE, 
            sep = "\t", 
            quote = FALSE)
```

```{r}
significant_taxa_WT <- combined_df_WT %>%
  filter(Sign. != "ns") %>%
  pull(taxa) %>%
  unique()

combined_df_sig_WT <- combined_df_WT %>%
  filter(taxa %in% significant_taxa_WT)
```

```{r}
qmp_colors <- c(
  "Transit_Time"      = "#c04545", 
  "Fecal_Mass"      = "#264653", 
  "Microbiota_Excretion"    = "#fdb747",  
  "Microbial_Load" =  "#609A98"  
)

combined_df_sig_WT$QMP <- as.factor(combined_df_sig_WT$QMP)

combined_df_sig_WT$fill_color <- ifelse(
  combined_df_sig_WT$Significance == TRUE,
  qmp_colors[as.character(combined_df_sig_WT$QMP)],
  "white"
)

ordered_taxa_WT <- combined_df_sig_WT %>%
  filter(QMP == "Microbial_Load") %>%
  arrange(desc(Rho)) %>%
  pull(taxa)

combined_df_sig_WT$taxa <- factor(combined_df_sig_WT$taxa, levels = ordered_taxa_WT)
```

Plot
```{r} 
pdf("Taxa_QMP_corr_WT_fold.pdf", width = 14, height=14)
ggplot(combined_df_sig_WT, aes(x = taxa, y = Rho)) +
  geom_point(
    aes(color = QMP, fill = fill_color),
    shape = 21,
    size = 4,
    stroke = 1.2
  ) +
  scale_color_manual(
    name = "",
    values = qmp_colors,
     labels = c(
    "Transit_Time"      = "GI Transit Time [min]",
    "Fecal_Mass"      = "Fecal Mass [g per GI Passage]",
    "Microbiota_Excretion"    = "Microbiota Excretion [16S rRNA GCN per GI Passage]",
    "Microbial_Load" = "Microbial Load [16S rRNA GCN/g per GI Passage]"
  )
  ) +
  scale_fill_identity() +
  labs(
    x = "",
    y = "Spearman Correlation Coefficient (Rho)"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = c(-0.2, 0.2), linetype = "dotted", color = "gray")+
  Settings_corr_all
dev.off()
```