---
title: "Suppl. Figures"
output: html_document
date: "2026-01-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Suppl. Figure 2 - Mouse body weight diet study 

## 1. Preparation
```{r}
weight <- weight_diet %>%
  pivot_longer(
    cols = c(before_adaption, baseline, cycle_1, regular_1, regular_2, cycle_2),
    names_to = "timepoint",
    values_to = "weight"
  )

weight$timepoint <- factor(
  weight$timepoint,
  levels = c("before_adaption", "baseline", "cycle_1", "regular_1", "regular_2", "cycle_2"),
  labels = c("baseline", "start intervention", "GI passage\nexperiment 1", "week 2", "week 3", "GI passage\nexperiment 2")
)
```

## 2. Plot
```{r}
pdf("Body_weight_diet.pdf", width = 10, height=7)
ggplot(weight, aes(x = timepoint, y = weight, group = cage, color = diet)) +
  geom_point(alpha = 0.6) +
  stat_summary(
    fun = mean,
    geom = "line",
    aes(group = diet),
    linewidth = 1.2   
  ) +
  scale_color_manual(
    values = c(
      "control"   = "#85C1E9",  
      "fasting"   = "#AF7AC5",  
      "high-fat"  = "#F39C12",  
      "high-fiber"= "#52BE80"   
    ),
    name = "Diet",
    labels = c("Control", "Fasting", "High Fat", "High Fiber")
  ) +
  labs(
    x = "",  
    y = "Mouse Weight [g]"   
  ) +
    scale_y_continuous(limits = c(15, 40)) +
  Settings_line_weight
dev.off()
```

##3. Statistics
```{r}
x <- weight$weight

shapiro.test(x) 
ad.test(x)
```

```{r}
stats1_weight_diet <- weight %>% group_by (timepoint) %>% pairwise_wilcox_test(weight ~ diet) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

stats2_weight_diet <- weight %>% group_by (diet) %>% pairwise_wilcox_test(weight ~ timepoint) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")
```

# Suppl. Figure 3 - QMPs Clock Disruption

## 1 Preparation
```{r}
QMP_clock <- metadata_QP_passage %>%
    filter(project == "TRF_study") %>%
    filter(diet != "TRF")
```

```{r}
df_two <- QMP_clock %>%
  filter(
    (project == "TRF_study" & clock_group == "normal") |   # 12h/12h
      project == "TRF_study" & clock_group == "disrupted"                    # 4h/4h
  ) %>%
  mutate(
    condition = case_when(
      inactivity == 1 & clock_group == "normal" ~ "12h/12h",
      clock_group == "disrupted"               ~ "4h/4h",
      TRUE ~ NA_character_
    ),
    condition = factor(condition, levels = c("12h/12h", "4h/4h")),
    genotype_group = factor(genotype_group, levels = c("WT", "IL10-/-"))
  ) %>%
  select(condition, genotype_group,
         Fecal_Mass, Microbial_Load, Microbiota_Excretion, Transit_Time) %>%
  tidyr::pivot_longer(
    cols = c(Fecal_Mass, Microbial_Load, Microbiota_Excretion, Transit_Time),
    names_to = "QMP",
    values_to = "value"
  )
```

## 2. Plot
```{r}
y_limits <- tibble::tribble(
  ~QMP,               ~ymin,   ~ymax,
  "Fecal_Mass",        0,       1.0,
  "Microbial_Load",    2.5e10,   6e11,
  "Microbiota_Excretion",        1e10,     2.5e11,
  "Transit_Time",   50,       350
)

df_two <- df_two %>%
  left_join(y_limits, by = "QMP")

pos <- position_dodge(width = 0.8)

df_two$QMP <- factor(df_two$QMP,
                     levels = c("Transit_Time", "Fecal_Mass", "Microbial_Load", "Microbiota_Excretion"))

pos <- position_dodge(width = 0.75)

p_two <- ggplot(df_two, aes(
  x = genotype_group, y = value, fill = condition, color = condition)) +
  geom_boxplot(
    lwd = 1.2,
    width = 0.55,
    outlier.shape = NA,
    alpha = 0.25,
    position = pos
  ) +
  geom_point(
    size = 2.4,
    stroke = 0.7,
    position = pos,
    alpha = 0.8
  ) +
  geom_point(aes(y = ymin), alpha = 0) +
  geom_point(aes(y = ymax), alpha = 0) +
  facet_wrap(~ QMP, ncol = 4, scales = "free_y") +
  scale_fill_manual(values = c(
    "SLT 1" = "#3498DB",
    "SLT 5" = "#7F8C8D"
  )) +
  scale_color_manual(values = c(
    "SLT 1" = "#3498DB",
    "SLT 5" = "#7F8C8D"
  )) +
  scale_x_discrete(
    labels = c(
      "WT" = "WT",
      "IL10-/-" = expression("IL-10"^"-/-"))) +
  labs(
    x = "",
    y = "",
    fill = "",
    color = "") +
  Settings_inactivity +
  theme(
    panel.spacing = unit(4, "lines"))

ggsave("QMPs_clockdisruption.pdf",
       p_two, width = 28, height = 10)

```

##3. Statistics

```{r}
summary_stats_clock_disruption <- QMP_clock %>%
  group_by(genotype_group, clock_group) %>%
  summarise(
    Fecal_Mass_mean = mean(Fecal_Mass, na.rm = TRUE),
    Fecal_Mass_sd   = sd(Fecal_Mass, na.rm = TRUE),

    Microbial_Load_mean = mean(Microbial_Load, na.rm = TRUE),
    Microbial_Load_sd   = sd(Microbial_Load, na.rm = TRUE),

    Microbiota_Excretion_mean = mean(Microbiota_Excretion, na.rm = TRUE),
    Microbiota_Excretion_sd   = sd(Microbiota_Excretion, na.rm = TRUE),

    Transit_Time_mean = mean(Transit_Time, na.rm = TRUE),
    Transit_Time_sd   = sd(Transit_Time, na.rm = TRUE)
  )
```


```{r}
x <- QMP_clock$Fecal_Mass

shapiro.test(x) 
ad.test(x)
```
--> normal distribution

```{r}
stats1_stool_clockTRF <- QMP_clock %>% group_by (genotype_group) %>% pairwise_t_test(Fecal_Mass ~ clock_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

stats2_stool_clockTRF <- QMP_clock %>% group_by (clock_group) %>% pairwise_t_test(Fecal_Mass ~ genotype_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")
```

```{r}
x <- QMP_clock$Microbial_Load

shapiro.test(x) 
ad.test(x)
```
--> non-normal distribution

```{r}
stats1_density_clockTRF <- QMP_clock %>% group_by (genotype_group) %>% pairwise_wilcox_test(Microbial_Load ~ clock_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

stats2_density_clockTRF <- QMP_clock %>% group_by (clock_group) %>% pairwise_wilcox_test(Microbial_Load ~ genotype_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")
```


```{r}
x <- QMP_clock$Microbiota_Excretion

shapiro.test(x) 
ad.test(x)
```
--> non-normal distribution

```{r}
stats1_load_clockTRF <- QMP_clock %>% group_by (genotype_group) %>% pairwise_wilcox_test(Microbiota_Excretion ~ clock_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

stats2_load_clockTRF <- QMP_clock %>% group_by (clock_group) %>% pairwise_wilcox_test(Microbiota_Excretion ~ genotype_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")
```

```{r}
x <- QMP_clock$Transit_Time

shapiro.test(x) 
ad.test(x)
```
--> non-normal distribution

```{r}
stats1_transit_clock <- QMP_clock %>% group_by (genotype_group) %>% pairwise_wilcox_test(Transit_Time ~ clock_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

stats2_transit_clock <- QMP_clock %>% group_by (clock_group) %>% pairwise_wilcox_test(Transit_Time ~ genotype_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")
```


# Suppl. Figure 4 - Mouse body weight TRF study 
## 1. Preparation
```{r}
weight_TRF_long <- left_join(TRF_QMP, weight_TRF, by = "mouse")
weight_TRF_long <- column_to_rownames(weight_TRF_long, var = "mouse")

weight_TRF_long <- weight_TRF_long %>%
  select(cage, genotype_trf, GI_experiment, diet, genotype_group)
```


## 2. Plot
```{r}
TRF_weight_box <- weight_TRF_long

TRF_weight_box$genotype_trf <- factor(TRF_weight_box$genotype_trf, levels = c("control_WT", "TRF_WT", "control_IL10-/-", "TRF_IL10-/-"))
facet_labels_TRF <- c("control" = "ad libitum", "TRF" = "TRF")

facet_labels_genotype <- c("WT" = "WT", "IL10ko" = expression("IL-10"^"-/-"))

TRF_weight_box$genotype_group <- factor(TRF_weight_box$genotype_group, levels = c("WT", "IL10-/-"))
facet_labels_TRF <- c("control" = "ad libitum", "TRF" = "TRF")

facet_labels_genotype <- c("WT" = "WT", "IL10ko" = expression("IL-10"^"-/-"))

pdf("Weight_experiment_TRF.pdf", width = 9, height=6)
TRF_weight_box %>%
ggplot(aes(x= genotype_group,, 
                 y=GI_experiment, 
                 fill = genotype_trf,
                 color = genotype_trf)) + 
  geom_boxplot(lwd=1.5, width=0.6, outlier.shape = NA,  alpha = 0.2) +
  geom_point(size = 2.5, fill = "#333333") +
  scale_color_manual(
    values = c(
      "control_WT"   = "#3498DB",  
      "TRF_WT"   = "#2471A3",  
      "control_IL10-/-"  = "#E74C3C",  
      "TRF_IL10-/-"= "#A93226"   
    )) +
   scale_fill_manual(
    values = c(
      "control_WT"   = "#3498DB",  
      "TRF_WT"   = "#2471A3",  
      "control_IL10-/-"  = "#E74C3C",  
      "TRF_IL10-/-"= "#A93226"   
    )) +
   scale_x_discrete(
    labels = c("WT" = "WT ", "IL10-/-" = expression("IL-10"^"-/-")),
    expand=c(0.5, 0
             )) +
labs(x = "", 
         y = paste0("<span style='font-size: 25pt'>Body Weight</span><br><span style='font-size: 20pt'>[g]</span>")) + 
  facet_wrap(~diet, labeller = labeller(diet = facet_labels_genotype), scales = "fixed") +
  coord_cartesian(xlim = c(2, 1)) +
  scale_y_continuous(limits = c(22, 37)) +
  Settings_TRF +
  theme(axis.title.x = ggtext::element_markdown(), axis.title.y = ggtext::element_markdown())
dev.off()
```


## 3. Statistics
```{r}
x <- TRF_weight_box$GI_experiment

shapiro.test(x) 
ad.test(x)
```
--> non-normal distribution

```{r}
stats1_weight_TRF <- TRF_weight_box %>% group_by (genotype_group) %>% pairwise_t_test(GI_experiment ~ diet) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

stats2_weight_TRF <- TRF_weight_box %>% group_by (diet) %>% pairwise_t_test(GI_experiment ~ genotype_group) %>% adjust_pvalue(method = "BH") %>% add_significance("p.adj")

```


```{r}
markers <- c("GI_experiment")

mean_TRF_weight <- TRF_weight_box %>%
  group_by(genotype_trf) %>%
  summarise(
    across(
      all_of(markers),
      list(
        mean = ~ mean(.x, na.rm = TRUE),
        sd   = ~ sd(.x, na.rm = TRUE)
      )
    ),
    .groups = "drop"
  )

for (m in markers) {
  mean_col <- paste0(m, "_mean")
  sd_col   <- paste0(m, "_sd")
  pm_col   <- paste0(m, "_mean_pm_sd")

  mean_TRF_weight[[pm_col]] <-
    paste0(
      round(mean_TRF_weight[[mean_col]], 4),
      " ± ",
      round(mean_TRF_weight[[sd_col]], 4)
    )
}

write.table(mean_TRF_weight, file = 'mean_TRF_weight.txt', sep = "\t",quote = FALSE, row.names = FALSE)
```

#Suppl. Figure 5A - Shannon Index TRF study
#Suppl. Figure 5B - PCoA Plot TRF study
#Suppl. Figure 7 - dbRDA diet & TRF study controlled for project
--> see Figure 6C 